Difference-in-Differences with Spatial Spillovers

### 2021-03-22 latex/slides/Spillover_slides.tex
Update slides for Environmental Brownbag

@@

    \begin{itemize}
        \item[1--] Using a potential outcomes framework, I decompose the difference-in-differences estimator into three parts:
        \begin{itemize}
            \item Direct Effect of Treatment, Spillover onto Treated Units, Spillover onto Control Units
        \end{itemize}

        \pause
        \item[2--] Analyze effectiveness of common ad-hoc adjustments done by researchers
        \begin{itemize}
            \item Show that an indicator for being close to treated units remove \textit{all bias} so long as the indicator contains all units affected by spillovers
            \item `Rings' are able to estimate spillover effects while still removing all bias
        \end{itemize}

        \pause
        \item[3--] Apply framework to Urban Economics
        \begin{itemize}
            \item Revisit \citet{Kline_Moretti_2014a} analysis of the Tennessee Valley Authority

            \item Discuss how framework can reconcile conflicting findings on effect of federal Empowerment Zones \citep{Busso_Gregory_Kline_2013,Neumark_Kolko_2010}
        \end{itemize}
    \end{itemize} 
\end{frame}

\begin{frame}{Contribution}

    {\footnotesize

    \underline{Difference-in-Differences Estimation with Spillovers}

@@

    \begin{itemize}
        \item I generalize the work by deriving the bias in terms of potential outcomes and show how to estimate treatment effects by removing spillover effects
    \end{itemize}

    \pause
    \underline{Potential Outcome Framework for Spillovers}

    \begin{citecolor}
        [\citet{Miguel_Kremer_2004}, \citet{Vazquez-Bare_2019}, \citet{Angelucci_DiMaro_2016}, \citet{Angrist_2014}]        
    \end{citecolor}

    \begin{itemize}
        \item These papers consider spillovers in network clusters. I focus on spatial setting
    \end{itemize}

    \pause
    \underline{Place-based Policies}

    \begin{citecolor}
        [\citet{Kline_Moretti_2014b}, \citet{Kline_Moretti_2014a}, \citet{Busso_Gregory_Kline_2013}, and \citet{Neumark_Kolko_2010}]
    \end{citecolor}

    \begin{itemize}
        \item I highlight the need to control for general equilibrium effects to properly estimate local effects of policies
    \end{itemize}
    }
\end{frame}

@@

    \vspace{10mm}
    This is the direct effect in the absense of exposure to spillovers.
\end{frame}

@@

\begin{frame}{What does Diff-in-Diff identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the diff-in-diff estimate as follows:

@@

\begin{frame}{Biased Estimate for $\tau_{\text{direct}}$}
    \[ 
        \mathbb{E}[\hat{\tau}_{\text{diff-in-diff}}] - \textcolor{green}{\tau_{\text{direct}}} = \textcolor{red}{\tau_{\text{spillover, treated}}} - \textcolor{purple}{\tau_{\text{spillover, control}}}    
    \]
\end{frame}

@@

@@

    \begin{itemize}
        \item The TVA was a large-scale federal investment started in 1934 that focused on improving manufacturing economy.

        \item The program focused on electrification through dams and transportation canals in order to improve the manufacturing economy

        \item Hundreds of dollars spent anually per person
    \end{itemize}
\end{frame}

@@

    \begin{itemize}
        \item Agricultural employees might be drawn to hire wages for new manufacturing jobs in Tennessee Valley (negative spillover on control units)

        \item Manufacturing jobs that would have been created in the control units in the absence of treatment might move to the Tennessee Valley (negative spillover on control units)

        \item Since manufacutring jobs pay higher than agricultural jobs, the above two spillover effects can cause changes in median family income
    \end{itemize}
\end{frame}

>>>>

\begin{frame}{This Paper}
    In this paper, I

    \begin{itemize}
        \item Present a potential outcomes framework to formalize spillover effects and evaluate ad-hoc adjustments done in the literature

        \item Propose an estimation strategy that improves on current practices by being more robust to spillovers

        \item Apply this framework to improve estimation of the local effect of place-based policies in Urban Economics 
    \end{itemize}

@@

    \begin{itemize}
        \item[1--] Formalize spillovers into a potential outcomes framework:

@@

        \begin{itemize}
            \vspace{2.5mm}
            \item I decompose the difference-in-differences estimator into three parts: Direct Effect of Treatment, Spillover onto Treated Units, Spillover onto Control Units

            \vspace{2.5mm}
            \item Show that an indicator for being close to treated units remove \textit{all bias} so long as the indicator contains all units affected by spillovers

            \vspace{2.5mm}
            \item `Rings' are able to estimate spillover effects while still removing all bias
        \end{itemize}

\begin{frame}{Outline}
    \begin{itemize}
        \item[2--] Apply framework to Urban Economics
        \begin{itemize}
            \item Revisit \begin{citecolor}\citet{Kline_Moretti_2014a}\end{citecolor} analysis of the Tennessee Valley Authority

            \begin{itemize}
                \vspace{2.5mm}
                \item The local effect estimate is contaminated by spillover effects to neighboring counties \begin{citecolor}\citep{Kline_Moretti_2014b}\end{citecolor}

                \vspace{2.5mm}
                \item Large scale manufacturing investment creates an `urban shadow' \begin{citecolor}\citep{Cuberes_Desmet_Rappaport_2021,Fujita_Krugman_Venables_2001}\end{citecolor}
            \end{itemize}

            \vspace{2.5mm}
            \item Discuss how framework can reconcile conflicting findings on effect of federal Empowerment Zones \begin{citecolor}\citep{Busso_Gregory_Kline_2013,Neumark_Kolko_2010}\end{citecolor}
        \end{itemize}
    \end{itemize} 
\end{frame}

@@

    \pause
    \vspace{2.5mm}
    Spillover Effects:

    \[
        \textcolor{red}{\tau_{\text{spillover, treated}}} \equiv \textcolor{red}{\mathbb{E} \left[ Y_{i1}(1, h_i(\vec{D})) - Y_{i1}(1, 0) \mid D_i = 1 \right]}
    \]

    \[ 
        \textcolor{purple}{\tau_{\text{spillover, control}}} = \textcolor{purple}{\mathbb{E} \left[ Y_{i1}(0, h_i(\vec{D})) - Y_{i1}(0, 0) \mid D_i = 0 \right]}
    \]
\end{frame}

@@

\begin{frame}{What does Difference-in-Differences identify?}
    With the parallel trends assumption and random assignment of $D_i$, I decompose the diff-in-diff estimate as follows:

@@

@@

    \onslide<3->{
        \textbf{Inidicator that captures all affected unit removes all bias}
    }
\end{frame}

@@

    \begin{itemize}
        \item The TVA was a large-scale federal investment started in 1934 that focused on improving manufacturing economy. (Hundreds of dollars spent anually per person)

        \item The program focused on large-scale dam construction that brought cheap wholesale electricity to the region
    \end{itemize}

    Research Question: What are the local effects of TVA on manufacturing and agricultural economies? Do these effects come at the cost of other counties?
\end{frame}

@@

    \begin{itemize}
        \item \textbf{Agriculture:} 
        \begin{itemize}
            \item Employees might be drawn to hire wages for new manufacturing jobs in Tennessee Valley (negative spillover on control units)
        \end{itemize}

        \item \textbf{Manufacturing:}
        \begin{itemize}
            \item Cheap electricity might be available to nearby counties (positive spillover on control units)

            \item Manufacturing jobs that would have been created in the control units in the absence of treatment might move to the Tennessee Valley (negative spillover on control units)
        \end{itemize}

        \item \textbf{Income:} 
        \begin{itemize}
            \item Since manufacutring jobs pay higher than agricultural jobs, the above two spillover effects can cause changes in median family income
        \end{itemize}
    \end{itemize}
\end{frame}
